---
title: "GenomicSelection"
author: "massaine"
date: "2022-06-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Getting started

```{r}
library(genomicMateSelectR)

blups<-readRDS(here::here("output","DRGBlupsGS.rds"))
blups %>% dplyr::select(Trait, fitAS) %>% unnest(fitAS)
blups <- blups[-c(4,6,8),]

blups1<-readRDS(here::here("output","DRGBlupsDisGS.rds"))
blups1 %>% dplyr::select(Trait, fitAS) %>% unnest(fitAS)
blups <- rbind(blups,blups1)

for (i in 1:nrow(blups)) {
blups$fitAS[[i]]$blups[[1]] <- blups$fitAS[[i]]$blups[[1]] %>% dplyr::rename(germplasmName = GID)
  }


```

## Kinship matrix GBS DART
```{r}
A<-readRDS(file=here::here("output","GAddmatrixGBSandDArTGS.rds"))  ## GBS DART
#A<-readRDS(file=here::here("output","GAddmatrixDArTGS.rds"))  ## GBS DART

A[1:5,1:5]
dim(A)
hist(A[lower.tri(A)], main='Off-diagonal values (lower-triangle) of the kinship matrix')
hist(diag(A), main='Diagonals of the kinship', breaks=20)

```

```{r}
blups %<>% dplyr::select(Trait,fitAS) %>% unnest(fitAS) %>% dplyr::select(Trait,blups) %>% 
     # need to rename the "blups" list to comply with the runCrossVal function
     dplyr::rename(TrainingData=blups) %>% 
     dplyr::select(Trait,TrainingData) %>% 
     # need also to remove phenotyped-but-not-genotyped lines
     # couldn't hurt to also subset the kinship to only phenotyped lines... would save RAM
     dplyr::mutate(TrainingData=map(TrainingData,
                             ~filter(.,germplasmName %in% rownames(A)) %>% 
                                  # rename the germplasmName column to GID
                                  dplyr::rename(GID=germplasmName)))
for(i in 1:nrow(blups)){
blups$TrainingData[[i]] <- blups$TrainingData[[i]][order(blups$TrainingData[[i]]$GID),]
}
blups
```

```{r}
# For fastest, lightest compute of accuracy, remove non-phenotyped from kinship
gids<-blups %>% unnest(TrainingData) %$% unique(GID)
A<-A[gids,gids]
A <- A[order(rownames(A)),order(rownames(A))]
dim(A)
A2 <- A2[gids, gids]
D2 <- D
#save(A,A2,blups,file="Blups_Kinship_Alltraits.RData")

load(here::here("data", "Blups_Kinship_Alltraits.RData"))
rm(A)
A2 <- readRDS(here::here("output", "GAddmatrixDArTGS.rds"))
A2 <- A2[gids, gids] %>%
  .[order(rownames(.)),order(rownames(.))]
D2 <- readRDS(here::here("output", "GDommatrixDArTGS.rds"))
D2 <- D2[gids,gids] %>% .[order(rownames(.)),order(rownames(.))]

dim(D2)
D2[1:8,1:8]
A2[1:8,1:8]

```

## Selection index (Verificar pesos)
```{r}
# I chose to remove MCMDS 
## our preliminary analysis showed it to have ~0 heritability in this dataset
## initial test of cross-val. showed the models do not fit
SIwts<-c(DMCg=15,
         logFRY=20,
         logFSY=20)
SIwts
```

# Execute cross-validation
```{r}

Clones <- blups$TrainingData %>% tail(n=4) %>% head(n=1) %>% .[[1]] %>% .$GID
library(genomicMateSelectR)
starttime<-proc.time()[3]
standardCV<-runCrossVal(blups=blups %>% tail(n=4),
                        modelType="AD",
                        selInd=F,
                        grms=list(A=A2[Clones,Clones], D=D2[Clones,Clones]),
                        nrepeats=2,nfolds=5,
                        gid="GID",seed=424242,
                        ncores = 2, nBLASthreads = 4)

```

## Save the results
```{r}
saveRDS(standardCV,file = here::here("output","standardCV_DARTxGBS.rds"))  ## GBSXDART
#saveRDS(standardCV,file = here::here("output","standardCV_DART.rds"))  ## DART

```

## plot results
```{r}
library(ggplot2)
standardCV %>% 
     unnest(accuracyEstOut) %>% 
     dplyr::select(repeats,id,predOf,Trait,Accuracy, predOf) %>% 
  mutate(Trait = factor(Trait, levels = c("DMCg", "logFSY", "logFRY", "PA",
                                          "StC", "Anthractnose", "ManchaBranca",
                                          "ManchaParda", "QueimaFolhas"))) %>% 
     ggplot(.,aes(x=Trait,y=Accuracy,fill=predOf)) + 
     geom_boxplot() + theme_bw()

standardCV %>% 
     unnest(accuracyEstOut) %>% 
     dplyr::select(repeats,id,predOf,Trait,Accuracy, predOf) %>% 
  group_by(Trait, predOf) %>% summarize(AccMean = mean(Accuracy, na.rm = TRUE))
standardCV2 %>% 
     unnest(accuracyEstOut) %>% 
     dplyr::select(repeats,id,predOf,Trait,Accuracy, predOf) %>% 
  group_by(Trait, predOf) %>% summarize(AccMean = mean(Accuracy, na.rm = TRUE))

```


### Set-up for the predictions

```{r}
blups<-readRDS(here::here("output","DRGBlupsGS.rds"))
blups %>% dplyr::select(Trait, fitAS) %>% unnest(fitAS)
blups <- blups[-c(4,6,8),]

blups1<-readRDS(here::here("output","DRGBlupsDisGS.rds"))
blups1 %>% dplyr::select(Trait, fitAS) %>% unnest(fitAS)
blups <- rbind(blups,blups1)

for (i in 1:nrow(blups)) {
blups$fitAS[[i]]$blups[[1]] <- blups$fitAS[[i]]$blups[[1]] %>% dplyr::rename(germplasmName = GID)
  }

A<-readRDS(file=here::here("output","GAddmatrixDArTGS.rds"))
D<-readRDS(file=here::here("output","GDommatrixDArTGS.rds"))
```

```{r}
blups %<>% dplyr::select(Trait,fitAS) %>% unnest(fitAS) %>% dplyr::select(Trait,blups) %>% 
     # need to rename the "blups" list to comply with the runCrossVal function
     dplyr::rename(TrainingData=blups) %>% 
     dplyr::select(Trait,TrainingData) %>% 
     # need also to remove phenotyped-but-not-genotyped lines
     # couldn't hurt to also subset the kinship to only phenotyped lines... would save RAM
     dplyr::mutate(TrainingData=map(TrainingData,
                             ~filter(.,germplasmName %in% rownames(A)) %>% 
                                  # rename the germplasmName column to GID
                                  dplyr::rename(GID=germplasmName)))
blupsRes <- readRDS(file = here::here("output", "DRGBlupsDisGS.rds"))
blupsRes %>% select(Trait, fitAS) %>% unnest(fitAS)
```

##Run genomic predictions
```{r}
gpreds<-runGenomicPredictions(modelType="AD",
                              selInd=F,
                              blups=blups,
                              grms=list(A=A2, D=D2),
                              ncores=3, nBLASthreads = 2)
```

## Extract GEBV

##We have a single-row tibble.
##To access a simple table listing GEBV for each trait and the selection index:
```{r}
gpreds
gpreds$gblups[[1]] %>% filter(GID %like% "BR-21") -> gpredC3
write.table(x = gpredC3, file = here::here("output", "C3GETGVPred.csv"), sep = "\t",
            quote = F)
write.table(x = gpreds[[1]], file = here::here("output", "GETGVPred.csv"), sep = "\t",
            quote = F)

```

##At this point, you can use the SELIND predictions directly to rank and select parents.
##Example: sort by SELIND and pick the top 10…
```{r}
gpreds$gblups[[1]] %>% 
     arrange(desc(SELIND)) %>% 
     slice(1:10)
```

##For more detailed output, including variance component estimates:
```{r}
gpreds$genomicPredOut[[1]]
```

```{r}
gpreds$genomicPredOut[[1]]$varcomps[[1]]
```

```{r}
saveRDS(gpreds,file = here::here("output","genomicPredictions.rds"))
```

